How Kids Code and How We Know:\\An Exploratory Study on the Scratch Repository

### 2016-06-13 Paper/paper.tex
Fixed minor error
De-anonymized

@@

\numberofauthors{1}
\author{
% 1st. author
\alignauthor
(anonymized submission)
%Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{1932 Wallamaloo Lane}\\
%       \affaddr{Wallamaloo, New Zealand}\\
%       \email{trovato@corporation.com}
% 2nd. author
%\alignauthor
%G.K.M. Tobin\titlenote{The secretary disavows
%any knowledge of this author's actions.}\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{P.O. Box 1212}\\
%       \affaddr{Dublin, Ohio 43017-6221}\\
%       \email{webmaster@marysville-ohio.com}
}

@@

\terms{terms}

@@

Several studies have shown that block-based languages are powerful as tools for teaching programming~\cite{meerbaum-salant_learning_2010, b._moskal_evaluating_2005,cooper_teaching_2003,price_comparing_2015}. Previous works involving static analysis of Scratch programs have evaluated the application of various programming concepts in Scratch projects \cite{Maloney_2008,moreno_automatic_2014}. Recent works have focused on bad programming practices within Scratch programs \cite{Meerbaum_habits_2011}, and automated quality assessment tools have been proposed for identifying code smells~\cite{fowler_refactoring:_1999}~and bad programming practices \cite{boe_hairball:_2013,moreno_automatic_2014}. A recent controlled experiment found that long scripts and code duplication decreases a novice programmer's ability to understand and modify Scratch programs \cite{hermans_2016}.

@@

Our study is based on data from the Scratch project repository. By scraping the list of recent projects,\footnote{\label{scratchpublic}\url{https://scratch.mit.edu/explore/projects/all/}} we have obtained \nPrograms~public Scratch projects and performed source code analysis on them. To the best of our knowledge, this is the first large-scale exploratory study of Scratch projects.

@@

\begin{description}
	\item [Sprites] Scratch code is organized by `sprites': two dimensional pictures each having their own associated code. Scratch allows users to bring their sprites to life in various ways, for example by moving them in the plane, having them say or think words or sentences via text balloons, but also by having them make sounds, grow, shrink and switch costumes. The Scratch program in Figure \ref{fig:ui} contains two one sprites, the cat, which is Scratch's default sprite and a piano. The code in Sprite1 will cause the cat to move right when the right arrow is pressed, and when the green flag is clicked it will continuously sense touching the piano.
	\item [Scripts]Sprites can have multiple code blocks, called scripts. The Scratch code in Figure \ref{fig:ui} has two distinct scripts, one started by clicking on the green flag and one by pressing the space bar. It is possible for a single sprite to have multiple scripts initiated by the same event. In that case, all scripts will be executed simultaneously.
	\item [Events] Scratch is \emph{event-driven}: all motions, sounds and changes in the looks of sprites are initiated by events called Hat blocks\footnote{\url{http://wiki.scratch.mit.edu/wiki/Hat_Block}}). The canonical event is the \texttt{when Green Flag clicked}, activated by clicking the green flag at the top of the user interface. In addition to the green flag, there are a number of other events possible, including key presses, mouse clicks and input from a computer's microphone or webcam. The Scratch code in Sprite1 in Figure \ref{fig:ui} contains two events: \texttt{when Green Flag clicked} and \texttt{when <right arrow> key pressed}, each with associated blocks.
	\item [Signals] Events within Scratch can be user generated too: users can broadcast a message, for example when two sprites touch each other, like in Figure \ref{fig:ui}. All other sprites can then react by using the \texttt{when I receive} Hat block. In Figure \ref{fig:ui}, Sprite1 broadcasts `bump' when the cat touches the Piano. 
	\item[Custom blocks] Scratch users can define their own blocks, which users can name themselves, called custom blocks. The creation of custom blocks in Scratch is the equivalent of defining procedures in other languages \cite{moreno_automatic_2014}. Because the term `procedures' is common in related work, we will refer to custom blocks as `procedures' in the remainder of this paper. Procedures can have input parameters of type string, number, and boolean. When a user defines a procedure, a new Hat block called \texttt{define} appears, which users can fill with the implementation of their block.
\end{description}

@@

We ran Kragle on March 2nd 2016 for 24 hours and, in that time, it obtained a little over 250,000 projects. Out of the \nPrograms, we failed to parse and further analyze 2,367 projects due to technical difficulties with the provided JSON files. Kragle, as well as all scraped projects and our analysis files are available.\footnote{\label{repo}\url{https://github.com/ScratchLover42/ICER-Data-Code} (temporary link to protect the anonymity of the authors)}

>>>>

\def\sharedaffiliation{%
\end{tabular}
\begin{tabular}{c}}

@@

\numberofauthors{2}
\author{
	\alignauthor Efthimia Aivaloglou\\      
	\email{e.aivaloglou@tudelft.nl}
	%
	\alignauthor Felienne Hermans\\     
	\email{f.f.j.hermans@tudelft.nl}
	%
	\sharedaffiliation
	\affaddr{Software Engineering Research Group}\\
	\affaddr{Delft University of Technology }   \\
	\affaddr{Mekelweg 4, 2628 CD, Delft, the Netherlands }
}

@@

%\terms{terms}

@@

Several studies have shown that block-based languages are powerful as tools for teaching programming~\cite{meerbaum-salant_learning_2010, b._moskal_evaluating_2005,cooper_teaching_2003,price_comparing_2015}. Previous works involving static analysis of Scratch programs have evaluated the application of various programming concepts in Scratch projects \cite{Maloney_2008,moreno_automatic_2014}. Recent works have focused on bad programming practices within Scratch programs \cite{Meerbaum_habits_2011}, and automated quality assessment tools have been proposed for identifying code smells~\cite{fowler_refactoring:_1999}~and bad programming practices \cite{boe_hairball:_2013,moreno_automatic_2014}. In a recent controlled experiment we found that long scripts and code duplication decreases a novice programmer's ability to understand and modify Scratch programs \cite{hermans_2016}.

@@

Our study is based on data from the Scratch project repository. By scraping the list of recent projects,\footnote{\label{scratchpublic}\url{https://scratch.mit.edu/explore/projects/all/}} we have obtained \nPrograms~public Scratch projects and performed source code analysis on them. To the best of our knowledge, this is the first large-scale exploratory study of Scratch programs.

@@

\begin{description}
	\item [Sprites] Scratch code is organized by `sprites': two dimensional pictures each having their own associated code. Scratch allows users to bring their sprites to life in various ways, for example by moving them in the plane, having them say or think words or sentences via text balloons, but also by having them make sounds, grow, shrink and switch costumes. The Scratch program in Figure \ref{fig:ui} contains two sprites, the cat, which is Scratch's default sprite and a piano. The code in Sprite1 will cause the cat to move right when the right arrow is pressed, and when the green flag is clicked it will continuously sense touching the piano.
	\item [Scripts]Sprites can have multiple code blocks, called scripts. The Scratch code in Figure \ref{fig:ui} has two distinct scripts, one started by clicking on the green flag and one by pressing the space bar. It is possible for a single sprite to have multiple scripts initiated by the same event. In that case, all scripts will be executed simultaneously.
	\item [Events] Scratch is \emph{event-driven}: all motions, sounds and changes in the looks of sprites are initiated by events called Hat blocks\footnote{\url{http://wiki.scratch.mit.edu/wiki/Hat_Block}}). The canonical event is the \texttt{when Green Flag clicked}, activated by clicking the green flag at the top of the user interface. In addition to the green flag, there are a number of other events possible, including key presses, mouse clicks and input from a computer's microphone or webcam. The Scratch code in Sprite1 in Figure \ref{fig:ui} contains two events: \texttt{when Green Flag clicked} and \texttt{when <right arrow> key pressed}, each with associated blocks.
	\item [Signals] Events within Scratch can be user generated too: users can broadcast a message, for example when two sprites touch each other, like in Figure \ref{fig:ui}. All other sprites can then react by using the \texttt{when I receive} Hat block. In Figure \ref{fig:ui}, Sprite1 broadcasts `bump' when the cat touches the Piano. 
	\item[Custom blocks] Scratch users can define their own blocks, which users can name themselves, called custom blocks. The creation of custom blocks in Scratch is the equivalent of defining procedures in other languages \cite{moreno_automatic_2014}. Because the term `procedures' is common in related work, we will refer to custom blocks as `procedures' in the remainder of this paper. Procedures can have input parameters of type string, number, and boolean. When a user defines a procedure, a new Hat block called \texttt{define} appears, which users can fill with the implementation of their block.
\end{description}

@@

We ran Kragle on March 2nd 2016 for 24 hours and, in that time, it obtained a little over 250,000 projects. Out of the \nPrograms, we failed to parse and further analyze 2,367 projects due to technical difficulties with the provided JSON files. Kragle, as well as all scraped projects and our analysis files are available.\footnote{\label{repo}\url{https://github.com/ScratchLover42/ICER-Data-Code} \todo{Felienne, what is the new address?} }
